Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-07-16 main.tex
triangulations

\begin{remark}
    Our main interest in this manuscript are finite triangulations of compact sets. 
    There is another reason why we insist on the triangulation being finite: 
    if we do not require our simplicial complexes to be finite,
    then a reasonable definition of simplicial complex would need additional topological conditions. 
    % 
    For example, if we modify our definition of simplicial complex to be infinite, 
    then any subset of Euclidean space would give rise to an ``infinite $0$-dimensional complex''. But this obviously does not reflect the topology. 
    %
    A more interesting such example is this:
    the Cantor set $\calC \subseteq [0,1]$ is a compact set 
    that underlies an ``infinite $0$-dimensional simplicial complex''.
    If we place $\calC$ on the x-axis of a 2D coordinate system 
    and connect each member of $\calC$ to $(0,1) \in \bbR^{2}$ via a straight line segment, 
    then the resulting set is still compact, even path-connected,  
    but has an ``infinite $1$-dimensional triangulation''.
    % 
    However, none of these infinite families are infinite simplicial complexes in the sense of geometric topology~\cite{spanier1995algebraic}. 
\end{remark}

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
Patches within triangulations of dimension two and three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures} are our main references for this section.

@@

We know that any manifold $\calM$ has got a topological boundary $\partial\calM$, possibly empty. 
If $\calM$ is $n$-dimensional, then the $\partial\calM$ is a topological manifold without boundary of dimension $n-1$.

The following special cases receive particular interest:
a \textit{$n$-ball triangulation} is any triangulation of a topological (closed) $n$-ball,
and we sometimes call this a \textit{$n$-disk} triangulation.
An \textit{$n$-sphere triangulation} is any triangulation of a topological $n$-sphere.

We gather a few helpful properties of manifold triangulations. 
While the reader might deem them obvious, we nevertheless include proofs. 
We begin with the boundary.

@@

@@

    If $z_F$ is an interior point of $\calM$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \cap \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. % TODO: proofread
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices. 
    \todo{More details}

    If $F$ is contained in only one $n$-simplex, then it must be a face in the boundary of $\calM$. 
    If $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each face within $\partial \calM$ is contained in exactly one $n$-simplex. 
    \todo{More details}

@@

@@

Let $T$ be an $n$-dimensional simplex with some subsimplex $F$ We show that there exists a hyperplane of codimension one that intersects $T$ only at the $F$ and such that $T$ is only on one side of the hyperplane.

\begin{lemma}\label{lemma:oppositesubsimplex}
    Let $v_0, v_1, \dots, v_n \in \bbR^n$ be the vertices of a simplex $T$.
    Let $F, F'$ be the two subsimplices of $T$ with respective sets of vertices $v_0,\dots,v_k$ and $v_{k+1},\dots,v_n$.
    Then $T$ lies on one side of the affine subspace
    \[
        A := v_0 + \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \} +  \operatorname{lin}\{ v_n - v_{k+1}, \dots, v_n - v_{n-1} \}.
    \]
    Moreover, $F = T \cap A$.
\end{lemma}
\begin{proof}
    Evidently, $A$ has dimension $n-1$ and $F \cap A \cap T$.
    Let us now suppose that $x \in A \cap T$.
    For some real coefficients $\lambda_1,\dots,\lambda_{n-1} \in \bbR$, it holds that 
    \begin{align}
        x 
        &
        = 
        v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n-1} \lambda_{i} ( v_n - v_i )
        \\&
        = 
        \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right) v_0 
        + 
        \sum_{i=1  }^{k  } \lambda_{i} v_i 
        + 
        \left( \sum_{i=k+1}^{n-1} \lambda_{i} \right) v_n
        + 
        \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i 
        .
    \end{align}
    We know $x$ can be written as a unique convex combination of the vertices of $T$.
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are non-positive. 
    If $x \notin F$, then at least one of these coefficients must be negative. 
    But then their sum must be negative, which is a contradiction. 
    The desired statement follows.         
\end{proof}

>>>>

\begin{remark}
    Our main interest in this manuscript are finite triangulations of compact sets. 
    There is another reason why we insist on the triangulation being finite: 
    if we do not require our simplicial complexes to be finite,
    then a reasonable definition of simplicial complex would need additional topological conditions. 
    % 
    For example, if we modify our definition of simplicial complex to be infinite, 
    then any subset of Euclidean space would give rise to an ``infinite $0$-dimensional complex''. But this obviously does not reflect the topology. 
    %
    A more interesting such example is this:
    the Cantor set $\calC \subseteq [0,1]$ is a compact set 
    that underlies an ``infinite $0$-dimensional simplicial complex''.
    If we place $\calC$ on the x-axis of a 2D coordinate system 
    and connect each member of $\calC$ to $(0,1) \in \bbR^{2}$ via a straight line segment, 
    then the resulting set is still compact, even path-connected,  
    but has an ``infinite $1$-dimensional triangulation''.
    % 
    However, none of these infinite families are infinite simplicial complexes in the sense of geometric topology~\cite{lee2011topological}. 
\end{remark}

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
Patches within triangulations of dimension two and three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures} are our main references for this section. 
We also refer to Lee's monograph~\cite{lee2011topological} for any further background on manifolds.

@@

The following special cases receive particular interest:
an \textit{$n$-ball triangulation} is any triangulation of a topological (closed) $n$-ball,
and we sometimes call this a \textit{$n$-disk} triangulation.
An \textit{$n$-sphere triangulation} is any triangulation of a topological $n$-sphere. 
\\

We know that any manifold $\calM$ has got a topological boundary $\partial\calM$, possibly empty. 
If $\calM$ is $n$-dimensional, then the $\partial\calM$ is a topological manifold without boundary of dimension $n-1$. 
We gather a few helpful observations on how these notions relate to triangulations.
While the reader might deem them obvious, we nevertheless include proofs.

@@

    \item 
    We prove an auxiliary result. 
    Suppose that $Y$ is a topological space homeomorphic to a sphere of dimension $m$ and that $X \subseteq Y$ is homeomorphic to a sphere of dimension $m-1$, where $m \geq 1$. 
    The embedding of $X$ into $Y$ cannot be onto~\cite[Theorem~2.55]{lee2011topological}, and so there exists $y \in Y \setminus X$. 
    Using the homeomorphism from $Y$ into the $m$-sphere, which maps $y$ to the northpole without loss of generality, and a stereographic projection, 
    we see that $Y \setminus \{y\}$ is homeomorphic to $R^{m}$.
    By the Jordan-Brouwer separation theorem~\cite{}, $Y \setminus (X \cup \{y\})$ has got two connected components,
    and thus $Y \setminus X$ has got two connected components.

@@

    Consider the case that $z_F$ lies on the boundary of $\calM$. 
    If $F$ is contained in at least two distinct $n$-simplices of $\calT$,
    then we can construct an embedded $n$-dimensional open ball that contains $z_F$.
    But that contradicts $z_F$ being on the boundary. 
    So any boundary face can only be contained in one single $n$-simplex.

    If $z_F$ is an interior point of $\calM$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere of dimension $n-1$. 
    At the same time, $X = F \cap \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere of dimension $n-2$. % TODO: proofread
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by the auxiliary result above, $K = 2$. 
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
%     But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
%     which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
%     If an inner face $F$ were contained in only one $n$-simplex,
%     then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
%     So each inner face is contained in exactly two $n$-simplices. 
%     \todo{More details}

%     If $F$ is contained in only one $n$-simplex, then it must be a face in the boundary of $\calM$. 
%     If $F$ were contained in at least two $n$-simplices,
%     then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
%     So each face within $\partial \calM$ is contained in exactly one $n$-simplex. 
%     \todo{More details}
%     {\color{blue}
%     The doubling $M_{D}$ of the manifold $M$ is obtained by gluing $M$ to a copy of itself along the boundary.
%     The triangulation $\calT$ can be doubled to get a triangulation of the doubled manifold.
%     Any boundary face is an interior face of the doubled manifold two $n$-simplices of the doubled triangulation.
%     So each face within $\partial \calM$ is contained in exactly one $n$-simplex of $\calT$. 
%     }

@@

Suppose that any $n$-dimensional simplex $T$ with some subsimplex $F$.
We want to find a hyperplane of codimension one that intersects $T$ only at the $F$ and such that $T$ is only on one side of the hyperplane.

\begin{lemma}\label{lemma:oppositesubsimplex}
    Let $v_0, v_1, \dots, v_n \in \bbR^n$ be the vertices of a simplex $T$.
    Let $F, F'$ be the two subsimplices of $T$ with respective sets of vertices $v_0,\dots,v_k$ and $v_{k+1},\dots,v_n$.
    Then $T$ lies on one side of the affine subspace
    \[
        A := v_0 + \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \} +  \operatorname{lin}\{ v_n - v_{k+1}, \dots, v_n - v_{n-1} \}.
    \]
    Moreover, $F = T \cap A$.
\end{lemma}
\begin{proof}
    Evidently, $A$ has dimension $n-1$ and $F \cap A \cap T$.
    Let us now suppose that $x \in A \cap T$.
    For some real coefficients $\lambda_1,\dots,\lambda_{n-1} \in \bbR$, it holds that 
    \begin{align}
        x &= 
        v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n-1} \lambda_{i} ( v_n - v_i )
        \\&
        = 
        \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right) v_0 
        + 
        \sum_{i=1  }^{k  } \lambda_{i} v_i 
        + 
        \left( \sum_{i=k+1}^{n-1} \lambda_{i} \right) v_n
        + 
        \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i 
        .
    \end{align}
    We know $x$ can be written as a unique convex combination of the vertices of $T$.
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are non-positive. 
    If $x \notin F$, then at least one of these coefficients must be negative. 
    But then their sum must be negative, which is a contradiction. 
    The desired statement follows.         
\end{proof}

\begin{theorem}
    Let $\calT$ be an $n$-dimensional triangulation and let $S \in \calT$ be a simplex of dimension $k < n$.
    Let $T \in \patch_{\calT}(S)$ be of dimension $n$ and let $M := |\patch_{\calT}(S)| \setminus T$. 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \varphi : T \rightarrow M
    \end{align*}
    satisfying
    \begin{align*}
        \Lip( \varphi,      T ) \leq \Cvier{n}{k},
        \quad 
        \Lip( \varphi^{-1}, M ) \leq \Cfive{n}{k},
    \end{align*}
    where $\Cvier{n}{k} \leq ...$ and $\Cfive{n}{k} \leq ...$
\end{theorem}

@@
